---
title: "ASSIGNMENT-1"
author: "SAI TEJA KANNEGANTI"
date: "28 August 2016"
output: word_document
---
DSA 5103 : INTELLIGENT DATA ANALYTICS
Required Packages for this assignment :
```{r}
  library (moments) # Contains functions skewness and kurtosis
  library (plyr)    # To load dataframe baseball
  library(datasets) # To load quakes data from datasets package
  library(reshape)   #contains replace function
```
PROBLEM 1 : Vectors using R

PROBLEM 1(a):
```{r}
x <- c(3,12,6,-5,0,8,15,1,-10,7) # To concatinate 10 numerical values in x.
```
Problem 1(b):
```{r}
y <- seq(min(x),max(x),length=10) #Finding minimum and maximum values of x and splitting interval into 10 numbers and assigning to y.
y
```
problem 1(c)
For x
```{r}
sum(x) 
mean(x) 
sd(x) 
var(x) 
mad(x) # To find Mean Absolute Deviation.
quantile (x, probs=c(0.0,0.25,0.50,0.75,1.0)) # Displays value of random variable at these fractions.
quantile (x, probs=c(0,0.20,0.40,0.60,0.80,1.0))
```
For y
```{r}
sum(y) 
mean(y) 
sd(y) 
var(y) 
mad(y) 
quantile (y, probs=c(0.0,0.25,0.50,0.75,1.0))
quantile (y, probs=c(0.0,0.20,0.40,0.60,0.80,1.0)) 
```
quartiles are values of random variable that divide into 4 groups.
quintiles are values of random variable that divide into 5 groups.
problem 1(d):
```{r}
z<- sample (x,7, replace = TRUE) # Sampling random variable x,to choose 7 elements with replacement.
z
```
problem 1(e):
```{r}
skewness(x) 
kurtosis(x)# moments is package, skewness & kurtosis are funtions to determine skewness and kurtosis.
```
Skewness and kurtosis can also be determined in e1701 package, but this package is not available in R version 3.3.1
problem 1(f)
```{r}
t.test(x,y) # t-test of 2 independent vectors with some default parameters.
```
As p value is greater than 0.05 and 0 lies in the confidence interval, hypothesis is true,i.e. true difference between means is 0. So the difference of means is insignificant.
problem 1(g)
```{r}
sort(x) 
t.test(x, sort (x),paired=TRUE) # paired t-test between x and sort(x).
```
Since we are performing paired t-test between x and sort(x),mean of the differences is 0.
problem 1(h)
```{r}
n <- x[x<0] # n is a vector having negative valus of x.
n
```
problem 1(I)
```{r}
x <- x[!x %in% n]  # !x indicates not x,!x includes n,so x doesnot include n.
x
```
PROBLEM 2

problem 2 (a):
```{r}
college <- read.csv("E:/Masters/IDA/Assignments/Assignment 1/college.csv", header = TRUE, sep = ",") # read.csv() function to read tablular data & create a data frame. 
```
problem 2(b):
```{r}
rownames (college) <- college [,1] # Assigning first column(university name) as row names of dataframe.

View (college) # To View dataframe.

college <- college [,-1] # Removing first column of the dataframe.
```
problem 2(c)
1.
```{r}
summary(college) #Provides numerical summary of datavariables in college.
```
 2.
```{r}
pairs(college[,1:10]) # pairs is used to produce a matrix of scatter plots.
```
Scatterplot of 1 to 10 columns of dataframe college.
3.
```{r}
plot(college$Private, college$Outstate,main="Outstate Students in Private Universities",xlab = "Private", ylab = "Outstate")
```
Side by side boxplots of outstate students Vs private universities.
4.
```{r}
Elite <- rep ("NO", nrow(college)) # Replicates NO to all rows in college dataframe.
Elite [college$Top10perc >50] <- "Yes" # Elite contains universities having Top10perc greater than 50
Elite <- as.factor (Elite) # as.factor is used to encode a vector as a factor(1 for no,2 for yes)
college <- data.frame(college ,Elite) # To create Elite data variable in college dataframe
```
5.
```{r}
summary (Elite)
```
There are 78 Elite universities in given college list.
6.
```{r}
plot(college$Elite,college$Outstate, main="Outstate studets in Elite Universities",ylab="Outstate",xlab="Elite")
```
Side by side boxplots of outstate students in Elite universities.
7.
```{r}
par(mfrow=c(3,2)) # used for multiple number of plots simultaniously (r*c).
hist(college$Apps,breaks=9) # Histogram for no of applications(Apps) with 9 breaks
hist(college$Accept,breaks=11)
hist(college$Enroll,breaks=14)
hist(college$Outstate,breaks=9)
hist(college$Books,breaks=9)
hist(college$Grad.Rate,breaks=14)
```
Some histograms with differing numbers of bins for few of the quantitative variables.
Problem 3 : Manipulating data into data frames

Problem 3(a):
```{r}
data(baseball) # data loads specified data sets.
```
problem 3(b)
```{r}
baseball$sf[baseball$year< 1954] <- 0  # equating value of sf to 0,when year in baseball dataframe is less than 1954.
baseball$hbp[is.na(baseball$hbp)] <- 0 # Equating all missing values(NA) in hbp to 0.
baseball <- baseball[c(baseball$ab>=50), ] #Baseball contains rows of players whose ab >= 50(deletes rows of players with ab < 50) & all columns remains unchanged.
```
problem 3(c)
```{r}
obp <-((baseball$h+baseball$bb+baseball$hbp)/(baseball$ab+baseball$bb+baseball$hbp+baseball$sf)) 
baseball <- data.frame(baseball,obp) # combining obp variable(column) to baseball dataframe.
```
problem 3(d)
```{r}
Result <- baseball[order(-obp) , ] # Result is a dataframe containing rows sorted as per obp(-obp indicates decreasing order)
Result <- Result[1:5, ] # Considering only 1 to 5 rows.
Result[3:22] <- list(NULL) # Removing 3 to 22 columns.
head(Result)
```
Result contains player name, year and on base percentage of top 5 records.
problem 4

problem 4(a)
```{r}
data(quakes) # to load quakes from datasets package.
```
problem 4 (b)
```{r}
plot(quakes$depth,quakes$mag,main= "Earth quake Magnitude Vs Earth quake depth",xlab= "depth",ylab= "magnitude") # To plot a graph, need to give x-axis,y-axis quantities, we use $ inorder to access particular column in dataframe, later main is used for title, xlab,ylab used for labels.
```
Plot between Earth quake magnitude and depth from surface
problem 4 (c)
```{r}
quakeAvgDepth <- aggregate(quakes$depth,by = list (quakes$mag),mean) #we are using aggregate function to calculate average depth for given sequencence of magnitude,since we are calculating average depth, function written is mean.
```
problem 4(d)
```{r}
colnames(quakeAvgDepth) <- c("Magnitude","Avg_depth") # To list column names of dataframe.(we can also use rename command).
```
Renamed variables to Magnitude, Avg depth.
problem 4 (e)
```{r}
plot(quakeAvgDepth$Avg_depth,quakeAvgDepth$Magnitude,main= "Magnitude Vs Average earth quake depth",xlab= "Average depth",ylab= "magnitude") # Used renamed variables of "quakeAvgDepth" data frame to plot graph.
```
Plot between Earth quake magnitude and Average depth from surface
problem 4(f)
From the graphs we can understand that:
1. Generally magnitude of earthquake increases with decrease in depth from earth surface.(From plot 4(e))
2. More number of earthquakes occur at lower magnitudes(from 4.0 to 5.0).(From plot 4(b)).













